---
title: "OWMA Weasel Model 2026"
output:
  pdf_document: default
  html_notebook: default
---

# Parameters

## Files

`working_directory [char]`: The working directory, set with `setwd()`.

`PA_obs_path [char]`: Path to presence-absence observations CSV data.

`PA_locs_path [char]`: Path to presence-absence location CSV data.

`PO_path [char]`: Path to presence-only CSV data.

`shp_path [char]`: Path to a shapefile file boundary this project.

`boundary_paths [vec(name = char)]` : A named vector of paths to boundaries to be used. These should be shapefiles, pointing to a file with the ".shp" extension. Names will be used to assign boundaries to models in `model_defs`.

`covar_paths [vec(name = char)]`: A named vector of paths to custom (non-WorldClim) covariates to be used. These should be raster files such as tiff images. Names represent the name of the variable represented in the file, ex: "Grassland" for a file called "Grass_NLCD2011.tif". *intSDM* models will use the file name without the extension internally, rather than the name provided in this vector.

## Config

`species [vec(char)]`: A vector of species to include in this analysis.

`max_spatial_error [int/float]`: The maximum spatial distance error (in meters) to include. Records with values higher than this will be filtered out.

`date_cutoff [char]`: The earliest date to include in this analysis (formatted as MM/DD/YYYYY). Records with values lower than this will be filtered out.

`reproject [bool]`: If rasters should be reprojected into the CRS of the species data.

`species_proj [int]`: The projection code to use for coordinates provided with species data. This variable must be an integer, not a string.

`pred_dims [vec(int, int)]`: Dimensions (X, Y) of the output prediction map(s).

`quiet_workflow [bool]`: If the *intSDM* workflow should be quieted (minimally text output).

`worldclim_res [int/float]`: The resolution of WorldClim covariates. This must be one of the options available from WorldClim, which are provided in minutes. This variable must be an integer or float, not a string.

`project [char]`: The name to be used for this project internally for *intSDM* processes.

`seed [int]`: The seed to use for this script, used for reproducibility.

`PO_class_by [char]`: The name of the field defining which subset of data each record belongs to. A single likelihood will be created for each subset. This value is most often dataset names or a characteristic of the data such as what method was used to collect it. If only one likelihood is needed, a field must be created that contains only a single value.

`PA_class_by [char]`: The name of the field defining which subset of data each record belongs to. A single likelihood will be created for each subset. This value is most often dataset names or a characteristic of the data such as what method was used to collect it. If only one likelihood is needed, a field must be created that contains only a single value.

## Model

`model_defs [list(name = list("species" = vec(char), "boundary" = char, "priors" = vec("range" = vec(num, num), "sigma" = vec(num, num)), "Custom" = vec(char), "WorldClim" = vec(char), "WorldCover" = vec(char)))]`: A nested list representing the variables included in each model to be built. The outermost list contains named lists with the name of the model. Model names are used to form file paths and must conform to the characters allowed in folder names for your system. These contain a list with six named elements, "species", "boundary", "priors", "Custom", "WorldClim", and "WorldCover" representing the species the model is for, the boundary to be used, the hyperparameters (priors), and the three sources for variables. The species value is used to filter datasets to only the species of interest; a model will be built independently for each species value included in this vector. The boundary value is used to find and add the area of interest for the model; the name provided here will be used to find a file path provided in `boundary_paths`. Priors specifies the two priors used for the model, range and sigma; this vector must use the names "range" and "sigma" when specifying these variables. All three covariate elements contain a vector of variable names that correspond to the names provided in `covar_paths` or the names of variables present in WorldClimand ESA WorldCover's datasets. If no variable from one source is used, an empty vector should still be added in that place.

The following are valid options for variables from WorldClim and WorldCover:

-   WorldClim: "tmin", "tmax", "tavg", "prec", "srad", "wind", "vapr," "elev".

-   WorldCover: "trees", "grassland", "shrubs", "cropland", "built", "bare", "snow", "water", "wetland", "mangroves", "moss".

Example of an appropriate object:

``` r
model_defs <- list(
  "Model1" = list(
    "species" = c("Long-tailed weasel"),
    "boundary" = "FullState_Boundary",
    "priors" = c("range" = c(100, 1), "sigma" = c(1, 0.5)),
    "WorldClim" = c("tavg"),
    "WorldCover" = c("trees", "grassland"),
    "Custom" = c("Slope")
  ),

  "Model2" = list(
    "species" = c("Long-tailed weasel", "Short-tailed weasel"),
    "boundary" = "PartialState_Boundary",
    "priors" = c("range" = c(69, 67), "sigma" = c(4, 20)),
    "WorldClim" = c("tavg","vapr"),
    "WorldCover" = c(),
    "Custom" = c()
  )
)
```

`model_save [list(name = c(char))]`: A named list representing the models that should have point data extracted and converted to raster file. Names correspond to the names of models defined in `model_defs`. Values are vectors of which species' raster should be extracted (note that each model produces one map per species in your project if all were indicated in `model_defs`). The underlying data is stored as points and conversion to a raster results in some (minimal) data loss. If you experience missing lines of data horizontally or vertically in the output, adjust `rast_error`.

`rast_error [c(numeric, numeric)]`: The difference between the output raster's dimensions and the dimensions specified in `pred_dims`. The value specified here will be subtracted from `pred_dims`. Making a slightly smaller output raster will prevent missing lines of data in the output. For a 400x400 prediction surface, an error of only 2 cells is required.

```{r}
#Files 
working_dir <- 'C:/Users/Tanner/Desktop/OWMA_weasel_OHonly2'
PA_obs_path <- 'C:/Users/Tanner/Dropbox/OH-IN weasel manuscript/Data/csv_copies/PA_obs.csv'
PA_locs_path <- 'C:/Users/Tanner/Dropbox/OH-IN weasel manuscript/Data/csv_copies/PA_locs.csv'
PO_path <- 'C:/Users/Tanner/Dropbox/OH-IN weasel manuscript/Data/csv_copies/PO.csv'
boundary_paths <- list(
  'OH-IN_Full' = 'C:/Users/Tanner/Desktop/OH_only_OWMA_weasel/OH_buffer.shp',
  'OH-IN_NorthOnly' = 'C:/Users/Tanner/Desktop/OH_only_OWMA_weasel/OH_buffer.shp'
)

#boundary_paths <- list(
#  'OH-IN_Full' = 'C:/Users/Tanner/Dropbox/OH-IN weasel manuscript/Data/OH-IN_5kmBuffer/OH-IN_5kmBuffer.shp',
#  'OH-IN_NorthOnly' = 'C:/Users/Tanner/Dropbox/OH-IN weasel manuscript/Data/OH-IN_5kmBuffer_NorthOnly/OH-IN_5kmBuffer_NorthOnly.shp'
#)

covar_paths <- list(
  'agriculture' = 'C:/Users/Tanner/Dropbox/OH-IN weasel manuscript/Data/rasters/agriculture.tif',
  'developed' = 'C:/Users/Tanner/Dropbox/OH-IN weasel manuscript/Data/rasters/developed.tif',
  'forest' = 'C:/Users/Tanner/Dropbox/OH-IN weasel manuscript/Data/rasters/forest.tif',
  'grassland' = 'C:/Users/Tanner/Dropbox/OH-IN weasel manuscript/Data/rasters/grassland.tif',
  'water' = 'C:/Users/Tanner/Dropbox/OH-IN weasel manuscript/Data/rasters/water.tif',
  'SWE' = 'C:/Users/Tanner/Dropbox/OH-IN weasel manuscript/Data/rasters/swe.tif',
  'precipitation' = 'C:/Users/Tanner/Dropbox/OH-IN weasel manuscript/Data/rasters/prcp.tif',
  'tmax' = 'C:/Users/Tanner/Dropbox/OH-IN weasel manuscript/Data/rasters/tmax.tif',
  'tmin' = 'C:/Users/Tanner/Dropbox/OH-IN weasel manuscript/Data/rasters/tmin.tif'
)

#Config
species <- c('Long-tailed', 'Short-tailed', 'Least')
max_spatial_error <- 5000
date_cutoff <- '1/1/2015'
reproject <- FALSE
species_proj <- 4326

pred_dims <- c(615, 556)
quiet_workflow <- TRUE
worldclim_res = 0.5
project = 'OWMA Weasels 2026'
seed <- 890 #Generated from numbergenerator.org/randomnumbergenerator/1-1000

PO_class_by <- 'dataType'
PA_class_by <- 'camType'

#Model Info
model_save <- list(
  #'Long_tailed' = c('Long-tailed'), 
  'Short_tailed' = c('Short-tailed')#, 
  #'Least' = c('Least')
)
rast_error <- c(2, 2)
```

```{r}
#Model Definitions
model_defs <- list(
  #Long-tailed best
  'Long_tailed' = list(
    'species' = c('Long-tailed'),
    'boundary' = 'OH-IN_Full',
    'priors' = c('range' = c(30, 0.1), 'sigma' = c(1, 0.1)),
    'WorldClim' = c(),
    'WorldCover' = c(),
    'Custom' = c('developed', 'forest', 'grassland', 'water','SWE', 'tmax')
  ),
#  'Short_tailed' = list(
#    'species' = c('Short-tailed'),
#    'boundary' = 'OH-IN_NorthOnly',
 #   'priors' = c('range' = c(30, 0.1), 'sigma' = c(1, 0.1)),
  #  'WorldClim' = c(),
  #'WorldCover' = c(),
  #  'Custom' = c('agriculture', 'developed', 'forest', 'grassland', ,'SWE', 'tmax')
  #),
  'Least' = list(
    'species' = c('Least'),
    'boundary' = 'OH-IN_Full',
    'priors' = c('range' = c(30, 0.1), 'sigma' = c(1, 0.1)),
    'WorldClim' = c(),
    'WorldCover' = c(),
    'Custom' = c('agriculture', 'developed', 'grassland', 'water','SWE', 'tmax')
  )
)
```

```{r}
#TEST MODELS

#Model Definitions
model_defs <- list(
  #Long-tailed best
  'Long_tailed' = list(
    'species' = c('Long-tailed'),
    'boundary' = 'OH-IN_Full',
    'priors' = c('range' = c(30, 0.1), 'sigma' = c(1, 0.1)),
    'WorldClim' = c(),
    'WorldCover' = c(),
    'Custom' = c('agriculture', 'developed', 'grassland', 'water', 'SWE', 'tmax')
  ),
  'Short_tailed' = list(
    'species' = c('Short-tailed'),
    'boundary' = 'OH-IN_NorthOnly',
    'priors' = c('range' = c(30, 0.1), 'sigma' = c(1, 0.1)),
    'WorldClim' = c(),
    'WorldCover' = c(),
    'Custom' = c('agriculture', 'developed', 'grassland','SWE', 'tmax')
  ),
  'Least' = list(
    'species' = c('Least'),
    'boundary' = 'OH-IN_Full',
    'priors' = c('range' = c(30, 0.1), 'sigma' = c(1, 0.1)),
    'WorldClim' = c(),
    'WorldCover' = c(),
    'Custom' = c('agriculture', 'developed', 'grassland', 'water','SWE', 'tmax')
  )
)
```

# Initialization

## Install and Import Packages

```{r message=FALSE, warning=FALSE}
#List of Packages Used
packages <- c('MCMCglmm','tidyverse','remotes','INLA','dplyr','lubridate','PointedSDMs','intSDM','sf','terra','ggplot2') #'stringr'

#Check if Installed, Install if Not -- Import Regardless
for (pkg in packages) {
  if (!(require(pkg, character.only = TRUE))) {
    if (pkg == 'INLA') {
      install.packages(
        'INLA',
        repos = c(
          getOption("repos"),
          INLA = "https://inla.r-inla-download.org/R/stable"
        ), 
        dep = TRUE
      )
    } else if (pkg == 'ggregplot') {
      remotes::install_github("gfalbery/ggregplot")
    } else {
      install.packages(pkg)
    }
    
    library(pkg, character.only = TRUE)
  }
}

remove(packages, pkg)
```

## Set Seed and Working Directory

```{r}
#Set Seed
set.seed(seed)

#Set Working Directory
setwd(working_dir)
```

# Data Formatting

## Read in Datasets

```{r}
PA_obs <- read.csv(file = PA_obs_path) 
PA_locs <- read.csv(file = PA_locs_path) 
PO <- read.csv(file = PO_path)
```

```{r}
#Fix Column Types
PO$date <- lubridate::mdy(PO$date)
PA_obs$date <- lubridate::mdy(PA_obs$date)
PA_obs$date2 <- lubridate::mdy(PA_obs$date2)
PA_locs$setDate <- lubridate::mdy(PA_locs$setDate)
PA_locs$retrieveDate <- lubridate::mdy(PA_locs$retrieveDate)

#Create Generalized ID
## Accounts for reused names between projects
## Seperate with "::" incase it needs split later
PA_obs$ID <- paste0(PA_obs$dataSource, "::", PA_obs$camera_site, '::', PA_obs$deployment)
PA_locs$ID <- paste0(PA_locs$dataSource, "::", PA_locs$camera_site, '::', PA_locs$deployment)
```

## Filter

```{r}
#Presence-only Data
PO <- PO |>
  filter(
    !is.na(latitude), # Rem NULL latitude
    !is.na(longitude), # Rem NULL longitude
    species %in% .env$species, # Rem "Unknown" records
    positionalError_meters <= max_spatial_error, #Rem Large Spatial Error
    !(is.na(positionalError_meters)), #Rem Unknown Error
    !('S' %in% positionalError_meters), #Rem Unknown Error
    date >= mdy(date_cutoff) #Remove dates before Cutoff
  )

#Presence-Absence Data
PA_obs <- PA_obs |>
  filter(
    species %in% .env$species, #Rem "Unknown" records
  ) 
  
PA_locs <- PA_locs |>
  filter(
    locFuzz == 0, #Rem fuzzed locations
    positionalError_meters < max_spatial_error, #Rem Large Spatial Error
    !(is.na(positionalError_meters)), #Rem Unknown Error
  )
```

## Copy Columns for Presence-Absence Data

```{r}
#Copy Locational Data Columns to Presence-Absence Observations
PA_obs <- merge(
  x = PA_obs, 
  y = PA_locs[,c('ID',PA_class_by,'latitude','longitude')], 
  by = 'ID', 
  all.x = TRUE
)
```

## Save Copy of Datasets After Filter

```{r}

dir.create(file.path(working_dir, project))

write.csv(PO, file = file.path(working_dir, project, 'PO_filter.csv'))

write.csv(PA_obs, file.path(working_dir, project, 'PA_filter.csv'))
```

## Create Occupancy Matrix

```{r}
#Convert to Long Format
PA_long <- PA_obs |>
  group_by(ID, date, species) |>
  summarize(count = n()) |>
  ungroup()
```

```{r}
#Create Columns for All Dates
##Get List of All Dates in Range
dates <- as.character(
  seq(
    min(PA_locs$setDate), 
    max(PA_locs$retrieveDate), 
    by = 'days'
  )
)

##Get List of Sites
sites <- unique(PA_locs$ID)

##Initialize DF
PA_occu <- data.frame(matrix(ncol = 8))

##Add Columns and Row Value
colnames(PA_occu) <- c('ID','date','day_num','species','detection','latitude','longitude', PA_class_by)
```

```{r}
#Create Matrix
for (i in 1:nrow(PA_locs)) {
  for (sp in species) {
    ##Get List of Days Camera was Active
    dep_days <- as.character(
      seq(
        PA_locs[i,'setDate'],
        PA_locs[i,'retrieveDate'],
        by = 'days'
        )
    )
    
    ##Get Days with Observations for ID + Species
    det_days <- filter(
      PA_long, PA_long$ID == PA_locs[i, 'ID'] & PA_long$species == sp)$date
    
    ##Create Boolean Detection Values
    if (length(det_days) > 0) {
      occu_days <- ifelse(dep_days %in% det_days, 1, 0)
    } else {
      occu_days <- rep(0, length(dep_days))
    }

    ##Create DF
    x <- data.frame(
      ID = rep(PA_locs[i, 'ID'][[1]], length(dep_days)),
      date = dep_days,
      day_num = 1:length(dep_days),
      species = rep(sp, length(dep_days)),
      detection = occu_days,
      latitude = rep(PA_locs[i, 'latitude'][[1]], length(dep_days)),
      longitude = rep(PA_locs[i, 'longitude'][[1]], length(dep_days))
    )
    
    x[[PA_class_by]] <- rep(PA_locs[i, PA_class_by][[1]], length(dep_days))
    
    ##Add DF to DF
    PA_occu <- bind_rows(PA_occu, x)
  }
}

#Remove Empty First Row
PA_occu <- PA_occu[-c(1),] 

#Save Data
write.csv(PA_occu, file.path(working_dir, project, 'PA_occu_filter.csv'))

remove(x, occu_days, dep_days, det_days, i, sp)
```

## Format Spatial Data

```{r}
#Convert Presence Data to ST
PO <- st_as_sf(x = PO, coords = c('longitude','latitude'), crs = crs(paste0("+init=epsg:",species_proj)))
PA_occu <- st_as_sf(x = PA_occu, coords = c('longitude','latitude'), crs = crs(paste0("+init=epsg:",species_proj)))

#Add Boundaries
boundaries <- list()
for (boundary in names(boundary_paths)) {
  shp <- st_read(boundary_paths[[boundary]])
  
  ##Save to Object
  boundaries[[boundary]] <- shp
}

##Change Reference to WGS84
if (crs(shp) != crs("+init=epsg:4326")) {
  shp <- st_transform(shp, crs("+init=epsg:4326"))
}
if (crs(PO) != crs("+init=epsg:4326")) {
  PO <- st_transform(PO, crs("+init=epsg:4326"))
}
if (crs(PA_occu) != crs("+init=epsg:4326")) {
  PA_occu <- st_transform(PO, crs("+init=epsg:4326"))
}

##Assign Names from Path Object
names(boundaries) <- names(boundary_paths)


#Load Rasters
rasters <- list()
index <- 1

#Get Rasters as Rast Objects
for (var in covar_paths) {
  ##Add Obj to Raster List
  rasters[[names(covar_paths)[index]]] <- scale(terra::rast(var))
  names(rasters[[index]]) <- names(covar_paths)[[index]]
  
  ##Reproject
  if (reproject == TRUE) {
    cat('\n----- Reprojecting Raster-----\n')
    rasters[[index]] <- terra::project(x = rasters[[index]], crs(paste0("+init=epsg:", species_proj)))
  }
  ##Update Index
  index <- index + 1
  
}
```

## Raster Covariance Matrix

```{r}
i <- 1
for (r in rasters) {
  if (i == 1) {
    rast_stack <- c(r)
  } else {
    if (crs(r) != crs(rasters[[1]])) {
      r <- project(r, rasters[[1]])
    }
    
    #Resample to Same Extent
    r <- terra::resample(x = r, y = rasters[[1]], method = 'average')
    rast_stack <- c(rast_stack, r)

  }
  #Update Index
  i <- i + 1
}

#Create Covariance Matrix
cov_matrix <- data.frame(terra::layerCor(x = rast_stack, fun = 'cov', na.rm = TRUE))
cor_matrix <- data.frame(terra::layerCor(x = rast_stack, fun = 'cor', na.rm = TRUE))

#Save CSVs
write.csv(cov_matrix, file.path(working_dir, project, 'raster_covariance.csv'))
write.csv(cor_matrix, file.path(working_dir, project, 'raster_correlation.csv'))
```

# *intSDM* Model Workflow

```{r message=FALSE}
#Create Main Folder for Files
dir.create(
  path = file.path(working_dir, project), 
  showWarnings = FALSE
)
dir.create(
  path = file.path(working_dir, project, 'Model_Files'), 
  showWarnings = FALSE
)

#Loop through Model Definitions
i <- 1
for (model in model_defs) {
  skip <- FALSE
  
  #Check if Model Was Already Run
  ##Folder Exists
  if (dir.exists(file.path(working_dir, project, 'Model_Files', names(model_defs)[[i]]))) {
    skip <- TRUE
    
    ##Don't Skip if No Folders in Model Dir
    if (length(list.dirs(file.path(working_dir, project, 'Model_Files', names(model_defs)[[i]]), recursive = FALSE)) == 0) {
      skip <- FALSE
    }
    
    ##Check Number of Files in Each Folder
    for (folder in list.dirs(file.path(working_dir, project, 'Model_Files', names(model_defs)[[i]]), recursive = FALSE)) {
      ##If Any != 4, Don't Skip
      if (length(list.files(folder)) != 4) {
        skip <- FALSE
      }
    }
  }
      
  #Skip Model if Indicated
  if (skip == TRUE) {
    cat('\n----- Results for model: "', names(model_defs)[[i]], '" already exist. Skipping to next model. -----\n', sep = '')
    i <- i + 1
    next
  }
  
  #Start Workflow
  cat('\n-------------------- Starting Workflow... --------------------\n')
  cat('Model Name:', names(model_defs)[[i]], '\n')
  workflow <- startWorkflow(
    Projection =  "+proj=longlat +ellps=WGS84",
    Species = model_defs[[i]][['species']],
    saveOptions = list(projectName = project), 
    Save = FALSE,
    Quiet = quiet_workflow
  )

  ##Add Boundary
  cat('\n----- Adding Boundary... -----\n')
  workflow$addArea(Object = boundaries[[model_defs[[i]][['boundary']]]])
  
  ##Set Model Workflow Outputs
  cat('\n----- Setting Workflow Outputs... -----\n')
  workflow$workflowOutput(c('Model', 'Maps', 'Cross-validation'))

  #Add Occurence Data
  ##Presence-Only
  cat('\n----- Adding Presence-Only Data... -----\n')
  for (cls in unique(PO[[PO_class_by]])) {
    
    if (is.na(cls)) {
      next
    ### Skip adding if no records for this class
    } else {
      if (nrow(filter(filter(PO, !!sym(PO_class_by) == cls), species %in% model_defs[[i]][['species']])) == 0) {
        next
      }
    }
    
    
    workflow$addStructured(
      dataStructured = filter(PO, !!sym(PO_class_by) == cls),
      datasetType = 'PO',
      datasetName = paste0('PA__', gsub(' ', '_', cls)),
      speciesName = 'species',
      generateAbsences = TRUE
    )
  }
  
  ##Presence-Absence
  cat('\n----- Adding Presence-Absence Data... -----\n')
  for (cls in unique(PA_obs[[PA_class_by]])) {
    
    if (is.na(cls)) {
      next
    }
    
    workflow$addStructured(
      dataStructured = filter(PA_occu, !!sym(PA_class_by) == cls),
      datasetType = 'PA',
      datasetName = paste0('PA__', gsub(' ', '_', cls)),
      speciesName = 'species',
      responseName = 'detection',
      trialsName ='day_num'
    )
  }
  
  #Add Custom Covariates
  if (length(model$Custom) > 0) {
    cat('\n----- Adding Custom Covariates... -----\n')
    index = 1
    for (var in model$Custom) {
      cat('Adding Variable:', model$Custom[[index]], '\n')
      workflow$addCovariates(Object = rasters[[var]], Function = scale)
      
      index <- index + 1
    }
  } else {
    cat('\nNo Custom Covariates Added.\n')
  }

  #Add WorldClim Covariates
  if (length(model$WorldClim) > 0) {
    cat('\n----- Adding WorldClim Covariates... -----\n')
    index <- 1
    for (var in model$WorldClim) {
      cat('Adding Variable:', model$WorldClim[[index]], '\n')
      workflow$addCovariates(
        worldClim = var, 
        res = worldclim_res, 
        Function = scale
        )
      index <- index + 1
    }
  } else {
    cat('\nNo WorldClim Covariates Added.\n')
  }
  
  #Add WorldCover Covariates
  if (length(model$WorldCover) > 0) {
    cat('\n----- Adding WorldCover Covariates... -----\n')
    index <- 1
    for (var in model$WorldCover) {
      cat('Adding Variable:', model$WorldCover[[index]], '\n')
      workflow$addCovariates(
        landCover = var,
        Function = scale
        )
      index <- index + 1
    }
  } else {
    cat('\nNo WorldCover Covariates Added.\n')
  }

  #Add Mesh
  cat('\n----- Adding Mesh... -----\n')
  workflow$addMesh(
    cutoff = 0.2 * 5,
    max.edge = c(0.1, 0.24) * 80,
    offset = c(0.1, 0.4) * 100
  )
  
  #Specify Priors
  cat('\n----- Specifying Priors... -----\n')
  workflow$specifySpatial(
    prior.range = c(model_defs[[i]][['priors']][['range1']], model_defs[[i]][['priors']][['range2']]), #c(30, 0.1),
    prior.sigma = c(model_defs[[i]][['priors']][['sigma1']], model_defs[[i]][['priors']][['sigma2']])  #c(1, 0.1)
  )
  
  workflow$specifyPriors(
    effectNames = 'Intercept', 
    Mean = 0, 
    Precision = 1
  )
  
  #Cross Validation
  cat('\n----- Setting Cross Validating Info... -----\n')
  workflow$crossValidation(Method = 'Loo')
  
  #Create Model
  cat('\n----- Running Model... -----\n')
  model_data <- sdmWorkflow(
    workflow, 
    inlaOptions = list(
      num.threads = 1,
      control.inla = list(
        int.strategy = 'ccd',
        h = 1e-4,
        cmin = 0,
        control.vb=list(
          enable = FALSE
        )
      ),
      control.compute = list(
        return.marginals.predictor = TRUE
        ),
        safe = TRUE,
        verbose = !(quiet_workflow),
        inla.mode = 'experimental'
      ),
    predictionDim = pred_dims,
    ipointsOptions = list(method = 'direct')
  )
  ##Create Folder for Current Model
  dir.create(path = file.path(working_dir, project, 'Model_Files', names(model_defs)[[i]]), showWarnings = FALSE)
  
  #Copy RDS Files to New Folder
  cat('\n----- Saving Files... -----\n')
  #Save Data to RDS File
  for (sp in names(model_data)) {
    ##Create Folder for Current Model
    dir.create(path = file.path(working_dir, project, 'Model_Files', names(model_defs)[[i]], sp))
    for (rds in names(model_data[[sp]])) {
      saveRDS(model_data[[sp]][[rds]], file = file.path(working_dir, project, 'Model_Files', names(model_defs)[[i]], sp, paste0(rds, '.rds')))
    }
  }
  
  #Update Index
  i <- i + 1
  
  #Attempt to Free Some Memory
  gc(full = TRUE, verbose = FALSE)
}

#Clean Up Temporary Objects
remove(i, index, skip, cls, var, f, model)
```

# Model Statistics

```{r}
#Get a List of Folders to Look In
RDS_files <- list(
  c(), #paths
  c(), #model name
  c()  #species
)

for (model in names(model_defs)) {
  for (sp in model_defs[[model]][['species']]) {
  
    ##Get Path of File
    path <- file.path(working_dir, project, 'Model_Files', model, gsub(' ', '_', sp), 'Model.rds')
    
    ##Add to File List
    RDS_files[[1]] <- append(RDS_files[[1]], path)
    RDS_files[[2]] <- append(RDS_files[[2]], model)
    RDS_files[[3]] <- append(RDS_files[[3]], sp)
  }
}

#Read Each File and Get Statistics
stats_df = c()
for (i in 1:length(RDS_files[[1]])) {
  ##Check that File Exists
  if (!(file.exists(RDS_files[[1]][[i]]))) {
    warning(paste0('The file ', RDS_files[[1]][[i]], ' does not exist. Skipping...'))
    cat('\n')
    next
  }
  
  ##Read File
  RDS_data <- readRDS(RDS_files[[1]][[i]])
  
  stats <- c(
    
    ### General Info
    'model' = c(RDS_files[[2]][[i]]),
    'species' = c(RDS_files[[3]][[i]]),
    'covariates' = c(paste(RDS_data$spatCovs$name, collapse = ', ')),
    
    ### WAIC
    'WAIC' = c(RDS_data$waic$waic),
    'WAIC_p.eff' = c(RDS_data$waic$p.eff),
    
    ### DIC
    'DIC' = c(RDS_data$dic$dic),
    'DIC_p.eff' = c(RDS_data$dic$p.eff),
    'DIC_mean.deviance' = c(RDS_data$dic$mean.deviance),
    'DIC_deviance.mean' = c(RDS_data$dic$deviance.mean),
    
    ### Log marginal-likelihood
    'log.marginal.likelihood' = c(RDS_data$mlik[[1]]),
    
    ### RSS
    'RSS' = sum(RDS_data$residuals$deviance.residuals ** 2)
  )
  
  ##Fix Excel Display Issue
  if (startsWith(RDS_files[[2]][[i]],'-')) {
    stats[['model']] <- paste0("'", RDS_files[[2]][[i]])
  }
  
  ##Append to Dataset
  if (length(stats_df) == 0) {
    stats_df <- data.frame(t(data.frame(stats)))
  } else {
    stats_df <- rbind(stats_df, data.frame(t(data.frame(stats))))
  }
}

#Write Data to CSV
write.csv(x = stats_df, file = file.path(working_dir, project, 'Model_stats.csv'))

#Clean Up
remove(sp, i, stats, RDS_data, RDS_files)
```

# Save Rasters

```{r}
for (model in names(model_save)) {
  for (sp in model_save[[model]]) {
    #Read RDS Data
    RDS_file <- file.path(working_dir, project, 'Model_Files', model, gsub(' ', '_', sp), 'Predictions.rds')
    RDS_data <- readRDS(RDS_file)
    RDS_data <- RDS_data$predictions
    class(RDS_data) <- c('sf', 'data.frame') #IDK if this is needed
    
    #Convert Points to Raster
    preds <- terra::vect(RDS_data)
    rast <- terra::rast(
      preds, 
      ncols = pred_dims[1] - rast_error[[1]], 
      nrows = pred_dims[2] - rast_error[[2]]
    )
    mean_rast <- raster::rasterize(preds, rast, field = 'mean')
    std_rast <- raster::rasterize(preds, rast, field = 'sd')
  
    #Save Raster
    raster::writeRaster(x = mean_rast, filename = file.path(working_dir, project, paste0(model, '__', gsub(' ', '_', sp), '__predict.tif')), overwrite = TRUE)
    raster::writeRaster(x = std_rast, filename = file.path(working_dir, project, paste0(model, '__', gsub(' ', '_', sp), '__STD.tif')), overwrite = TRUE)
    
  }
}

#Clean Up
remove(sp, model, preds, rast, mean_rast, std_rast)
```

# Effect Size Data

```{r}
model_data <- list()
effsz_data <- data.frame()

for (model in names(model_save)) {
  model_data[[model]] <- list()
  
  for (sp in model_save[[model]]) {
    
    #Read RDS File
    RDS_file <- file.path(working_dir, project, 'Model_Files', model, gsub(' ', '_', sp), 'Model.rds')
    RDS_data <- readRDS(RDS_file)
    
    #Add to Model Object
    model_data[[model]][[sp]] <- RDS_data
    
    #Pull Table of Stats
    effsz_df <- model_data[[model]][[sp]][['summary.fixed']]
    
    #Fix Variable Column
    effsz_df[['variable']] <- rownames(effsz_df)
    rownames(effsz_df) <- 1:nrow(effsz_df)
    
    #Copy Model + Species Info
    effsz_df[['model']] <- model
    effsz_df[['species']] <- sp
    
    #Add to Full DataFrame
    effsz_data <- rbind(effsz_data, effsz_df)
  }
}

#Get Signifigance Info
effsz_data[['sig']] <- ifelse(effsz_data[['0.025quant']] > 0 | effsz_data[['0.975quant']] < 0, 1, 0)

#Save To CSV
write.csv(effsz_data, file = file.path(working_dir, project, 'effect_sizes.csv'))
```

## Plots

Currently, this is only made to handle one species per model in `model_save`. This will need to be updated if there are multiple models for a single species (i.e., if Model1 and Model2 are both for long-tailed weasels).

```{r}
#Order of Variables + Species, Change When Running with New Data
## Leave as empty list to use default order in DF
var_order = rev(names(covar_paths))
sp_order = c('Long-tailed', 'Short-tailed', 'Least')

#Filter Out Intercepts
effsz_data_plt <- effsz_data[!(grepl('intercept', effsz_data$variable)), ]

#Change Signifigance to Discrete Values for Color
effsz_data_plt[['sig']] <- as.character(effsz_data_plt[['sig']])

#Change Variable + Species Display Order
if (length(var_order) != 0) {
  effsz_data_plt[['variable']] <- factor(effsz_data_plt[['variable']], levels = var_order)
}

if (length(sp_order) != 0) {
  effsz_data_plt[['species']] <- factor(effsz_data_plt[['species']], levels = sp_order)
}

#Create Plot
effsz_plt <- ggplot(data = effsz_data_plt, aes(x=mean, y=variable, color=sig)) +
  #Theme + Styling
  theme_gray() +
  theme(panel.grid.major.x = element_blank(), panel.grid.minor.x = element_blank(), axis.title.y = element_blank(), legend.position = 'none') +
  labs(x = 'Effect Size') +
  geom_vline(xintercept = 0, linetype = 'longdash') +
  
  #Main Plot Elements
  geom_point() + 
  geom_errorbar(aes(xmin = effsz_data_plt[['0.025quant']], xmax = effsz_data_plt[['0.975quant']], width = 0.3)) +
  facet_grid(~species, scales = 'free')  +
  scale_color_manual(values=c("black", "brown3"))

#Save Plot
ggsave(
  filename = file.path(working_dir, project, 'effect_size.png'),
  plot = effsz_plt,
  dpi = 600,
  width = 6,
  height = 4
)

effsz_plt
```
